### 2023-03-29 response-to-audit-report.md

+  :: POSIXTime
+  -> State
+makeState minTime =
+  let
+    accounts = AM.singleton (Role "", Token "" "") 30_000_000  -- Note that 30 ada exceeds min-UTxO for current protocol parameters.
+    choices = AM.empty
+    boundValues = AM.empty
+  in
+    State{..}
diff --git a/marlowe-test/src/Spec/Marlowe/Semantics.hs b/marlowe-test/src/Spec/Marlowe/Semantics.hs
index dfd167f53..a15face41 100644
--- a/marlowe-test/src/Spec/Marlowe/Semantics.hs
+++ b/marlowe-test/src/Spec/Marlowe/Semantics.hs
@@ -24,6 +24,7 @@ import qualified Spec.Marlowe.Semantics.Entropy (tests)
 import qualified Spec.Marlowe.Semantics.Functions (tests)
 import qualified Spec.Marlowe.Semantics.Golden (tests)

+
 -- | Run the tests.
 tests :: TestTree
 tests =
diff --git a/marlowe-test/src/Spec/Marlowe/Semantics/Arbitrary.hs b/marlowe-test/src/Spec/Marlowe/Semantics/Arbitrary.hs
index b05563b29..3ab7ab9df 100644
--- a/marlowe-test/src/Spec/Marlowe/Semantics/Arbitrary.hs
+++ b/marlowe-test/src/Spec/Marlowe/Semantics/Arbitrary.hs
@@ -22,7 +22,8 @@

 module Spec.Marlowe.Semantics.Arbitrary
   ( -- * Types
-    IsValid(..)
+    Context
+  , IsValid(..)
   , SemiArbitrary(..)
     -- * Generators
   , arbitraryAssocMap
@@ -30,6 +31,7 @@ module Spec.Marlowe.Semantics.Arbitrary
   , arbitraryContractWeighted
   , arbitraryFibonacci
   , arbitraryGoldenTransaction
+  , arbitraryNonnegativeInteger
   , arbitraryPositiveInteger
   , arbitraryTimeIntervalAround
   , arbitraryValidInput
@@ -54,7 +56,15 @@ import Control.Monad (replicateM)
 import Data.Function (on)
 import Data.List (nub, nubBy)
 import Language.Marlowe.Core.V1.Semantics
-  (TransactionInput(..), TransactionOutput(..), computeTransaction, evalObservation, evalValue)
+  ( Payment(..)
+  , TransactionError(..)
+  , TransactionInput(..)
+  , TransactionOutput(..)
+  , TransactionWarning(..)
+  , computeTransaction
+  , evalObservation
+  , evalValue
+  )
 import Language.Marlowe.Core.V1.Semantics.Types
   ( Accounts
   , Action(..)
@@ -67,6 +77,7 @@ import Language.Marlowe.Core.V1.Semantics.Types
   , Environment(..)
   , Input(..)
   , InputContent(..)
+  , IntervalError(..)
   , Observation(..)
   , Party(..)
   , Payee(..)
@@ -77,6 +88,7 @@ import Language.Marlowe.Core.V1.Semantics.Types
   , ValueId(..)
   , getAction
   )
+import Plutus.Script.Utils.Scripts (dataHash)
 import Plutus.V2.Ledger.Api
   ( Credential(..)
   , CurrencySymbol(..)
@@ -87,11 +99,13 @@ import Plutus.V2.Ledger.Api
   , ValidatorHash(..)
   , adaSymbol
   , adaToken
+  , toBuiltinData
   )
 import PlutusTx.Builtins (BuiltinByteString, appendByteString, lengthOfByteString, sliceByteString)
 import Spec.Marlowe.Semantics.Golden (GoldenTransaction, goldenContracts, goldenTransactions)
+import Spec.Marlowe.Semantics.Merkle (merkleizeInputs, shallowMerkleize)
 import Test.Tasty.QuickCheck
-  (Arbitrary(..), Gen, chooseInteger, elements, frequency, listOf, shrinkList, sized, suchThat, vectorOf)
+  (Arbitrary(..), Gen, chooseInteger, elements, frequency, listOf, oneof, shrinkList, sized, suchThat, vectorOf)

 import qualified Plutus.V2.Ledger.Api as Ledger (Address(..))
 import qualified PlutusTx.AssocMap as AM (Map, delete, empty, fromList, keys, toList)
@@ -880,11 +894,11 @@ goldenContract = (,) <$> elements goldenContracts <*> pure (State AM.empty AM.em

 instance Arbitrary Contract where
   arbitrary = frequency [(95, semiArbitrary =<< arbitrary), (5, fst <$> goldenContract)]
-  shrink (Pay a p t x c) = [Pay a' p t x c | a' <- shrink a] ++ [Pay a p' t x c | p' <- shrink p] ++ [Pay a p t' x c | t' <- shrink t] ++ [Pay a p t x' c | x' <- shrink x] ++ [Pay a p t x c' | c' <- shrink c]
-  shrink (If o x y) = [If o' x y | o' <- shrink o] ++ [If o x' y | x' <- shrink x] ++ [If o x y' | y' <- shrink y]
-  shrink (When a t c) = [When a' t c | a' <- shrink a] ++ [When a t' c | t' <- shrink t] ++ [When a t c' | c' <- shrink c]
-  shrink (Let v x c) = [Let v' x c | v' <- shrink v] ++ [Let v x' c | x' <- shrink x] ++ [Let v x c' | c' <- shrink c]
-  shrink (Assert o c) = [Assert o' c | o' <- shrink o] ++ [Assert o c' | c' <- shrink c]
+  shrink (Pay a p t x c) = [c] ++ [Pay a' p t x c | a' <- shrink a] ++ [Pay a p' t x c | p' <- shrink p] ++ [Pay a p t' x c | t' <- shrink t] ++ [Pay a p t x' c | x' <- shrink x] ++ [Pay a p t x c' | c' <- shrink c]
+  shrink (If o x y) = [x, y] ++ [If o' x y | o' <- shrink o] ++ [If o x' y | x' <- shrink x] ++ [If o x y' | y' <- shrink y]
+  shrink (When a t c) = [c] ++ [When a' t c | a' <- shrink a] ++ [When a t' c | t' <- shrink t] ++ [When a t c' | c' <- shrink c]
+  shrink (Let v x c) = [c] ++ [Let v' x c | v' <- shrink v] ++ [Let v x' c | x' <- shrink x] ++ [Let v x c' | c' <- shrink c]
+  shrink (Assert o c) = [c] ++ [Assert o' c | o' <- shrink o] ++ [Assert o c' | c' <- shrink c]
   shrink _ = []

@@ -1047,10 +1061,22 @@ instance SemiArbitrary InputContent where
 instance Arbitrary Input where
   arbitrary = semiArbitrary =<< arbitrary
   shrink (NormalInput i)         = NormalInput <$> shrink i
-  shrink (MerkleizedInput i b c) = [MerkleizedInput i' b c | i' <- shrink i]
+  shrink (MerkleizedInput i b c) =
+    [NormalInput i]
+      <> [MerkleizedInput i' b c | i' <- shrink i]
+      <> [MerkleizedInput i (dataHash $ toBuiltinData c) c' | c' <- shrink c]

 instance SemiArbitrary Input where
-  semiArbitrary context = NormalInput <$> semiArbitrary context
+  semiArbitrary context =
+    frequency
+      [
+        (9, NormalInput <$> semiArbitrary context)
+      , (1, do
+              input <- semiArbitrary context
+              contract <- semiArbitrary context
+              pure $ MerkleizedInput input (dataHash $ toBuiltinData contract) contract
+        )
+      ]

 instance Arbitrary TransactionInput where
@@ -1073,7 +1099,7 @@ arbitraryValidStep :: State                 -- ^ The state of the contract.
                    -> Gen TransactionInput  -- ^ Generator for a transaction input for a single step.
 arbitraryValidStep _ (When [] timeout _) =
   TransactionInput <$> arbitraryTimeIntervalAfter timeout <*> pure []
-arbitraryValidStep state@State{..} (When cases timeout _) =
+arbitraryValidStep state@State{..} contract@(When cases timeout _) =
   do
     let
       isEmptyChoice (Choice _ []) = True
@@ -1090,7 +1116,13 @@ arbitraryValidStep state@State{..} (When cases timeout _) =
                                          Bound lower upper <- elements bs
                                          IChoice n <$> chooseInteger (lower, upper)
                     Notify _        -> pure INotify
-             pure $ TransactionInput times [NormalInput i]
+             is <-
+               frequency
+                 [
+                   (9, pure [NormalInput i])
+                 , (1, pure [MerkleizedInput i (dataHash $ toBuiltinData contract) contract])
+                 ]
+             pure $ TransactionInput times is
 arbitraryValidStep State{minTime} contract =
 {-
   NOTE: Alternatively, if semantics should allow applying `[]` to a non-quiescent contract
@@ -1144,10 +1176,66 @@ arbitraryValidInputs state contract =

 -- | Generate an arbitrary golden transaction.
-arbitraryGoldenTransaction :: Gen GoldenTransaction
-arbitraryGoldenTransaction =
+arbitraryGoldenTransaction :: Bool -> Gen GoldenTransaction
+arbitraryGoldenTransaction allowMerkleization =
   do
+    let
+      perhapsMerkleize gt@(state, contract, input, _) =
+        let
+          (contract', continuations) = shallowMerkleize contract
+          input' = merkleizeInputs continuations state contract' input
+        in
+          case input' of
+            Nothing      -> pure gt
+            Just input'' -> frequency [(9, pure gt), (1, pure (state, contract', input'', computeTransaction input'' state contract'))]
     equalContractWeights <- frequency [(1, pure True), (5, pure False)]
-    if equalContractWeights
-      then elements =<< elements goldenTransactions
-      else elements $ concat goldenTransactions
+    (if allowMerkleization then perhapsMerkleize else pure)
+      =<< if equalContractWeights
+            then elements =<< elements goldenTransactions
+            else elements $ concat goldenTransactions
+
+
+instance Arbitrary Payment where
+  arbitrary = Payment <$> arbitrary <*> arbitrary <*> arbitrary <*> arbitrary
+
+
+instance Arbitrary IntervalError where
+  arbitrary =
+    oneof
+      [
+        InvalidInterval <$> arbitrary
+      , IntervalInPastError <$> arbitrary <*> arbitrary
+      ]
+
+
+instance Arbitrary TransactionError where
+  arbitrary =
+    oneof
+      [
+        pure TEAmbiguousTimeIntervalError
+      , pure TEApplyNoMatchError
+      , TEIntervalError <$> arbitrary
+      , pure TEUselessTransaction
+      , pure TEHashMismatch
+      ]
+
+
+instance Arbitrary TransactionWarning where
+  arbitrary =
+    oneof
+      [
+        TransactionNonPositiveDeposit <$> arbitrary <*> arbitrary <*> arbitrary <*> arbitrary
+      , TransactionNonPositivePay <$> arbitrary <*> arbitrary <*> arbitrary <*> arbitrary
+      , TransactionPartialPay <$> arbitrary <*> arbitrary <*> arbitrary <*> arbitrary <*> arbitrary
+      , TransactionShadowing <$> arbitrary <*> arbitrary <*> arbitrary
+      , pure TransactionAssertionFailed
+      ]
+
+
+instance Arbitrary TransactionOutput where
+  arbitrary =
+    oneof
+      [
+        TransactionOutput <$> arbitrary <*> arbitrary <*> arbitrary <*> arbitrary
+      , Error <$> arbitrary
+      ]
diff --git a/marlowe-test/src/Spec/Marlowe/Semantics/Compute.hs b/marlowe-test/src/Spec/Marlowe/Semantics/Compute.hs
index 402460ed2..7aa9ed803 100644
--- a/marlowe-test/src/Spec/Marlowe/Semantics/Compute.hs
+++ b/marlowe-test/src/Spec/Marlowe/Semantics/Compute.hs
@@ -13,6 +13,7 @@

 {-# LANGUAGE ConstraintKinds #-}
 {-# LANGUAGE FlexibleContexts #-}
+{-# LANGUAGE LambdaCase #-}
 {-# LANGUAGE RankNTypes #-}
 {-# LANGUAGE RecordWildCards #-}

@@ -31,6 +32,7 @@ import Control.Applicative (liftA2)
 import Control.Lens.Getter (Getter, to, view)
 import Control.Monad.Except (MonadError(throwError), unless, when)
 import Control.Monad.Reader (ReaderT(runReaderT))
+import Data.Bifunctor (second)
 import Data.Default (Default(..))
 import Data.Function (on)
 import Data.List (sort)
@@ -49,29 +51,34 @@ import Language.Marlowe.Core.V1.Semantics.Types
   ( AccountId
   , Accounts
   , Action(Choice, Deposit, Notify)
-  , Case
+  , Bound(..)
+  , Case(..)
   , ChoiceId
   , ChosenNum
   , Contract(..)
   , Environment(Environment)
-  , Input
+  , Input(..)
   , InputContent(IChoice, IDeposit, INotify)
   , IntervalError(IntervalInPastError, InvalidInterval)
-  , Observation
+  , Observation(TrueObs)
   , Payee(Party)
   , State(..)
   , TimeInterval
   , Token(..)
-  , Value
+  , Value(Constant)
   , ValueId
   , getAction
   , getInputContent
   )
 import Language.Marlowe.FindInputs (getAllInputs)
-import Plutus.V2.Ledger.Api (CurrencySymbol, POSIXTime(..), TokenName)
+import Plutus.Script.Utils.Scripts (datumHash)
+import Plutus.V2.Ledger.Api (CurrencySymbol, Datum(..), DatumHash(..), POSIXTime(..), TokenName, toBuiltinData)
 import Spec.Marlowe.Semantics.Arbitrary
-  ( SemiArbitrary(semiArbitrary)
+  ( Context
+  , SemiArbitrary(semiArbitrary)
   , arbitraryContractWeighted
+  , arbitraryGoldenTransaction
+  , arbitraryPositiveInteger
   , assertContractWeights
   , closeContractWeights
   , defaultContractWeights
@@ -85,7 +92,19 @@ import Spec.Marlowe.Semantics.Orphans ()
 import System.IO.Unsafe (unsafePerformIO)
 import Test.Tasty (TestTree, testGroup)
 import Test.Tasty.QuickCheck
-  (Arbitrary(..), Gen, Testable(property), discard, elements, forAll, forAllShrink, frequency, suchThat, testProperty)
+  ( Arbitrary(..)
+  , Gen
+  , Testable(property)
+  , chooseInteger
+  , discard
+  , elements
+  , forAll
+  , forAllShrink
+  , frequency
+  , shuffle
+  , suchThat
+  , testProperty
+  )

 import qualified PlutusTx.AssocMap as AM

@@ -161,18 +180,145 @@ arbitraryMarloweContext w =
 -- | Generate an arbitrary valid Marlowe transaction context.
 arbitraryValid :: Gen MarloweContext
 arbitraryValid =
-    do
-      mcContract <- arbitrary `suchThat` (/= Close)
-      (time, inputs') <-
-        case unsafePerformIO $ getAllInputs mcContract of
-          Right candidates -> elements candidates
-          Left _           -> discard
-      let
-        -- TODO: Generalize to arbitrary starting state.
-        mcState = State AM.empty AM.empty AM.empty time
-        mcInput = head inputs'
-        mcOutput = computeTransaction mcInput mcState mcContract
-      pure MarloweContext{..}
+  let
+    randomContext =
+      do
+        mcContract <- arbitrary `suchThat` (/= Close)
+        (time, inputs') <-
+          case unsafePerformIO $ getAllInputs mcContract of
+            Right candidates -> elements candidates
+            Left _           -> discard
+        let
+          -- TODO: Generalize to arbitrary starting state.
+          mcState = State AM.empty AM.empty AM.empty time
+          mcInput = head inputs'
+          mcOutput = computeTransaction mcInput mcState mcContract
+        pure MarloweContext{..}
+    goldenContext =
+      do
+        (mcState, mcContract, mcInput, mcOutput) <- arbitraryGoldenTransaction True
+        pure MarloweContext{..}
+  in
+    frequency [(1, randomContext), (2, goldenContext)]
+
+
+-- | Generate a simple valid transaction.
+simpleTransaction
+  :: Context
+  -> (TimeInterval -> Contract -> Gen Contract)
+  -> (State -> Gen State)
+  -> (TransactionInput -> Gen TransactionInput)
+  -> Gen MarloweContext
+simpleTransaction context modifyContract modifyState modifyInput =
+  do
+    mcState <- modifyState =<< semiArbitrary context
+    intervalStart <- (getPOSIXTime (minTime mcState) +) <$> arbitraryPositiveInteger
+    intervalEnd <- (intervalStart +) <$> arbitraryPositiveInteger
+    let
+      interval = (POSIXTime intervalStart, POSIXTime intervalEnd)
+    timeout <- (intervalEnd +) <$> arbitraryPositiveInteger
+    mcContract <- modifyContract interval . When [] (POSIXTime timeout) =<< semiArbitrary context
+    mcInput <- modifyInput $ TransactionInput interval []
+    let
+      mcOutput = computeTransaction mcInput mcState mcContract
+    pure MarloweContext{..}
+
+
+-- | Generate a simple payment.
+simplePayment :: Gen MarloweContext
+simplePayment =
+  do
+    context <- arbitrary
+    balance <- arbitraryPositiveInteger
+    payment <- chooseInteger (1, balance)
+    account <- semiArbitrary context
+    payee <- semiArbitrary context
+    token <- semiArbitrary context
+    let
+      modifyContract _ contract = pure $ Pay account (Party payee) token (Constant payment) contract
+      modifyState state =
+        do
+          accounts' <-
+            fmap AM.fromList . shuffle . AM.toList
+              . AM.insert (account, token) balance
+              . AM.delete (account, token)
+              $ accounts state
+          pure state {accounts = accounts'}
+    simpleTransaction context modifyContract modifyState pure
+
+
+-- | Generate a simple deposit.
+simpleDeposit :: Gen MarloweContext
+simpleDeposit =
+  do
+    context <- arbitrary
+    deposit <- arbitraryPositiveInteger
+    account <- semiArbitrary context
+    payee <- semiArbitrary context
+    token <- semiArbitrary context
+    let
+      modifyContract (_, POSIXTime intervalEnd) contract =
+        pure
+          $ When [Case (Deposit account payee token $ Constant deposit) contract]
+            (POSIXTime $ intervalEnd + 1) contract
+      modifyInput (TransactionInput interval _) = pure $ TransactionInput interval [NormalInput $ IDeposit account payee token deposit]
+    simpleTransaction context modifyContract pure modifyInput
+
+
